---
title: "Getting Started with lyricsLab"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with lyricsLab}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE
)
```


# Welcome to lyricsLab!

lyricsLab is an R package for analyzing song lyrics and Spotify listening history. This guide will help you get started in just a few minutes.

## Installation

```{r}
# install.packages("devtools")
devtools::install_github("Coxabc/lyricsLab")
```

## Setup

### 1. Get Genius API Token (for lyrics analysis)

**Step 1:** Go to https://genius.com/api-clients  
**Step 2:** Sign up or log in (free)  
**Step 3:** Click "New API Client" and fill in:

- App Name: lyricsLab
- App Website: http://localhost
- Redirect URI: (leave default)

**Step 4:** Copy your "Client Access Token"  
**Step 5:** Set it in R:

```{r}
Sys.setenv(GENIUS_API_TOKEN = "your_token_here")
```

### 2. Get Spotify API Credentials (for listening history)

**Step 1:** Go to https://developer.spotify.com/dashboard  
**Step 2:** Log in with your Spotify account  
**Step 3:** Click "Create app" and fill in:

- App name: lyricsLab
- App description: R package for music analysis
- Redirect URI: `http://127.0.0.1:1410/`

**Step 4:** Copy your Client ID and Client Secret  
**Step 5:** Set them in R:

```{r}
Sys.setenv(SPOTIFY_CLIENT_ID     = "your_client_id")
Sys.setenv(SPOTIFY_CLIENT_SECRET = "your_client_secret")
```

> **Note:** Spotify's Web API requires a **Spotify Premium** account.

## Quick start

```{r}
library(lyricsLab)

# 1. Authenticate with Spotify
spotify_auth()

# 2. Get lyrics for a single song
lyrics <- get_lyrics("Ed Sheeran", "Drive")

# 3. Run analysis
sentiment_score(lyrics)
complexity_score(lyrics)
analyze_emotions(lyrics)
theme_detection(lyrics)

# 4. Plot
plot_sentiment_arc(lyrics)
plot_emotions_web(lyrics)
plot_themes(lyrics)
```

## Example: analyze a playlist

```{r}
spotify_auth()

analysis <- analyze_playlist(
  "https://open.spotify.com/playlist/37i9dQZF1DXcBWIGoYBM5M",
  max_songs = 50
)

plot_sentiment(analysis)
plot_complexity(analysis)
plot_scatter(analysis)
plot_top_complex(analysis)

calculate_listening_diversity(analysis)
```

## Example: your listening habits

```{r}
spotify_auth()
history <- get_recently_played()

plot_listening_patterns(history, type = "hourly")
plot_listening_patterns(history, type = "daily")
plot_listening_patterns(history, type = "weekly")

calculate_listening_diversity(history)
```

## Function reference

### Authentication & Spotify data

| Function | Description |
|---|---|
| `spotify_auth()` | Authenticate with Spotify (opens browser) |
| `get_recently_played()` | Your 50 most recently played tracks |
| `get_top_tracks()` | Your top tracks over short, medium, or long term |
| `get_playlist_tracks()` | All tracks from a playlist URL or ID |

### Lyrics

| Function | Description |
|---|---|
| `get_lyrics(artist, song)` | Fetch lyrics from Genius |
| `get_multiple_lyrics(artist, songs)` | Batch fetch lyrics for multiple songs |
| `preview_lyrics(lyrics)` | Print metadata and first few lines |
| `view_lyrics(lyrics)` | Print full lyrics to console |

### Analysis

| Function | Description |
|---|---|
| `sentiment_score(lyrics)` | Single overall sentiment score (0–100) |
| `sentiment_arc(lyrics)` | Line-by-line sentiment scores and trajectory |
| `analyze_emotions(lyrics)` | 8 NRC emotions with counts and percentages |
| `positive_negative_ratio(lyrics)` | Ratio of positive to negative lines |
| `complexity_score(lyrics)` | Vocabulary richness, reading level, complexity score |
| `theme_detection(lyrics)` | Keyword-based theme detection across 23 categories |

### Visualization

| Function | Description |
|---|---|
| `plot_sentiment_arc(lyrics)` | Line chart of sentiment through the song |
| `plot_emotions_bar(lyrics)` | Bar chart of 8 NRC emotions |
| `plot_emotions_web(lyrics)` | Radar chart of 8 NRC emotions |
| `plot_themes(lyrics)` | Bar chart of top themes |
| `plot_listening_patterns(history)` | Hourly, daily, or weekly listening patterns |

### Playlist analysis

| Function | Description |
|---|---|
| `analyze_playlist(playlist_url)` | Fetch and analyze all songs in a playlist |
| `plot_sentiment(analysis)` | Sentiment distribution across playlist |
| `plot_complexity(analysis)` | Complexity distribution across playlist |
| `plot_scatter(analysis)` | Sentiment vs complexity scatter |
| `plot_top_complex(analysis)` | Top 10 most complex songs |
